### 2016-02-28 hans/bayes-tardis-luminosity-prediction.tex
Inserted summary of major conceptual changes and possible consequences re (a) fitting all bins at the same time via rescaling (b) combining low-luminosity bins using the new bin index $m$. Important also the question whether we should work with frequency or wavelength.  Some comments on detail but no systematic work.

\usepackage[colorinlistoftodos]{todonotes} % Use 'disable' to remove todos in final version
\newcommand{\fred}[1]{\todo[inline]{#1}} %
\newcommand{\checked}{\todo[color=green,noline]{\checkmark}} %
\newcommand{\checkedl}{\todo[color=brown]{\checkmark}} %

%% CHANGE THIS DATE AS YOU MODIFY THE FILE
\newcommand{\vdate}{160218}
\pagestyle{myheadings}
\markboth{Version \vdate} {Version \vdate}
\parindent=0mm
\def\pp{\hfill\par \vspace{\baselineskip}}
%% \def\pp{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{bm,amssymb,amsfonts,amsmath}  % bold math symbols, AMStex
\usepackage{graphicx}    % standard graphics
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\lleq}[1]{\label{#1} }
% TO REMOVE THE LABEL LETTERS FROM THE EQUATION DISPLAY, COMMENT OUT THIS LINE:
\renewcommand{\lleq}[1]{\label{#1} {\scriptstyle {\rm (#1)}} \hspace*{2ex} }
% \renewcommand{\labelenumi}{\arabic{section}.\arabic{subsection}.\arabic{enumi}}
% \renewcommand{\thesubsubsection}{\arabic{subsubsection}}

@@

\DeclareMathOperator{\GammaDist}{Gamma}
\newcommand{\npack}{n_p}
\newcommand{\lmax}{\ell_{\rm max}}

@@

\textbf{Abstract} TODO

@@

\subsection{Definitions of relevant quantities}

\begin{mtemize}
\item Spectrum of luminosities in channels (or ``bins'')
  $b=0,1,\ldots,B$ is given by the telescope
\item In the simplest case, the frequency of channel $b$ is given by
  $\nu_b = \nu_0 + b\,\delta\nu$, where $\delta\nu$ is
  the bin width. In the general case, the bins are not of equal
  size. Let's then consider $\nu_b$ as the left edge of bin $b$,
  and $\nu_{B+1}$ is the right edge of the last bin; i.e., the
  maximum frequency.
  \item We treat the results of a single TARDIS run as \textit{data}
  \fred{If in the end we want to compare to actual data from a
    telescope, I'd prefer to not call the TARDIS output ``data''. Why
    don't we just call it ``TARDIS output''?}  $D$, i.e.\ fixed
  results which we do not question but accept as given.  These
  ``data'' are the number of packets leaving the supernova, $\npack$,
  the luminosity $\bml$, and the frequencies $\bmnu$ for each packet
  \begin{align}
    \lleq{dfc}
    D = \{ \npack, \bml, \bmnu \} \,,
  \end{align}
  where
  \begin{align}
    \lleq{dfd}
    % \bml_b &= (\ell_{b,1},\ell_{b,2},\ldots,\ell_{b,n_b})
    \dim \bml = \dim \bmnu = \npack \,.
  \end{align}
  Given $D$, the number of samples in bin $b$ is
  \begin{align}
    \lleq{dff}
    n_b = \sum_{j=1}^{\npack} \mathbf{1}_{[\nu_b,\nu_{b+1}]}( \nu_j )
  \end{align}
  and similarly the luminosity is
  \begin{align}
    \lleq{dfe}
    \ell_b = \sum_{j=1}^{\npack}\ell_j \mathbf{1}_{[\nu_b,\nu_{b+1}]}( \nu_j ),
  \end{align}
  where $\mathbf{1}_V(x)$ is the indicator function requiring $x \in V$.
\end{mtemize}

@@

While we treat the TARDIS output as ``data'' as discussed above, a
repeat of the simulation with a different Monte Carlo seed would
evidently yield different ``data'' $D'$ even when all other simulation
parameters remain the same. Clearly, $n_b$ is a specific realization
of a stochastic variable $N_b$ governed by a model pdf
$p(N_b)$. \fred{In our general case, $\npack = \sum_b n_b$ is an instance of a
  random variable because some packets leave the supernova, some
  don't. I guess in most applications the total number of events would
  be chosen and not random. We are fortunate that this difference does
  not matter because we consider bins independently. } Similarly,
$\bml_b$ is a specific realization of a collection of stochastic
variables $\bmL_b =\{L_{b,1},\ldots,L_{b,\smN_b}\}$, where the number
$N_b$ of such variables is itself stochastic. The relevant pdf for
$\bmL_b$ therefore always presupposes knowledge of $N_b$, so we must
work with the conditional probability $p(\bmL_b\cond N_b)$. This
situation is captured mathematically as $N_b \sim p(N_b), \bmL_b \sim
p(\bmL_b\cond N_b)$ where $\sim$ means ``is by hypothesis distributed
in this model according to the pdf'', i.e.\
\begin{align*}
  (N_b,\bmL_b) &\ \sim\ p(N_b,\bmL_b\cond \bmphi_b,\lambda_b)
  = p(\bmL_b\cond N_b,\bmphi_b)\; p(N_b\cond \lambda_b),
\end{align*}
where $\bmphi_b$ is the set of parameters of the model pdf which
govern $\bmL_b$ and $\lambda_b$ is the parameter of the model pdf
governing $N_b$.  To simulate $(N_b,\bmL_b)$, we clearly first need to
generate one $N_b$, followed by the generation of $N$ packet
luminosities $\bmL_b$. This situation is termed a \textit{compound
  process}.

@@

\begin{mtemize}
\item We omit the dependence on supernova parameters $\bmtheta$ from
  the notation below and instead condition on the TARDIS output
  $\npack, \bml, \bmnu$ directly.

\item In the remainder of this paper, we assume that the packet
  multiplicities $N_b$ and total luminosities $\bmL_b$ in any given
  bin $b$ are statistically independent of those in other bins. \fred{We could motivate this in the usual way as the Poisson limit of the Binomial. The subtlety would then the be the random total number of packets (leaving the supernova).} This
  means that the joint probability over all bins factorises into
  individual probabilities for each bin,
  \begin{align}
    \lleq{bsi}
    p(N_0,\bmL_0\ldots,N_b,\bmL_b,\ldots,N_\smB,\bmL_\smB
    \cond \npack, \bml, \bmnu,\bmphi,\bmlambda,\hmod)
    &= \prod_{b=0}^B p(N_b,\bmL_b\cond n_b, \bml, \bmnu,\bmphi,\bmlambda,\hmod)
    % p(N_0,\ldots,N_b,\ldots,N_\smB \cond \bmtheta,\bmphi,\hmod)
    % &= \prod_{b=0}^B p(N_b\cond \bmtheta,\bmphi,\hmod)\\
    % p(L_0,\ldots,L_b,\ldots,L_\smB \cond N_0,\ldots,N_\smB,\bmtheta,\bmphi,\hmod)
    % &= \prod_{b=0}^B p(L_b\cond N_b,\bmtheta,\bmphi,\hmod)
  \end{align}

@@

\item We choose $N$ to follow a Poisson distribution \fred{this choice
    linked to limit of Multinomial, see above} with parameter
  $\lambda$
  \begin{align}
    \lleq{pmc}
    p(N\cond\lambda) &= \frac{e^{-\lambda}\lambda^N}{N!}
    \qquad N = 0,1,\ldots,\infty,\quad \lambda>0
  \end{align}

@@

  \item In general, the distribution of $L_j$ could be a of a
  different functional form in each bin. For simplicity, we assume it
  is identical but may depend on parameters that differ from bin to
  bin. In practice, the distribution of packets does not change
  abruptly with frequency. We therefore assume that any parameter
  needed to fully specify $p(L_j)$ is given as a continuous function
  of frequency and parameters $\bmphi$; i.e.,
  \begin{align}
    \lleq{pmdb}
    p(L_j) = p(L_j \cond \nu, \bmphi)
  \end{align}
  For example, assume $p(L_j)$ depends on one parameter $a$. If we fit
  separately in each bin, we would have $B+1$ fits and $B+1$
  parameters. In each such fit, there would be only few samples to
  constrain $a$ in each bin. Then instead of fitting $a$ separately in
  each bin, we assume $a = a(\nu, \bmphi)$. This greatly reduces the
  number of parameters if $B$ is large because a handful of
  parameters, $\bmphi$, can be inferred from the complete set of
  packets. Necessarily, $a$ is best determined for the frequency
  around which the density of packets is highest.

@@

\item Priors: metaparameter $a$ \fred{?}

@@

the present case is also the Laplace transform. \fred{On Wikipedia,
  the MGF is defined as $\int_0^\infty dL\,p(L\cond
  n,\bml)\,e^{+tL}$. Then there are less minus signs to worry about}

@@

The Gamma distribution  with shape parameter $\alpha$ and rate
parameter $\beta$ is defined as
\begin{align}
  \lleq{gma}
  \GammaDist(x | \alpha, \beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} x^{\alpha-1} e^{-\beta x} .
\end{align}
The support of $\GammaDist(x \cond \alpha, \beta)$ is $[0, \infty]$
whereas the luminosity of real packets is bounded from above and has a
peak close to but not at the boundary. \fred{refer to plot of real
  luminosity packets in some frequency bin} In order to describe the
luminosity distribution of real packets through $\GammaDist(\alpha,
\beta)$, we have to transform the luminosities. To this end, we first
take the largest luminosity from the entire frequency range and
increase it slightly to avoid numerical difficulties at the end point
\begin{align}
  \lleq{gme}
  \lmax \equiv (1+\epsilon) \times \max_j \ell_j,
\end{align}
where $ \epsilon=10^{-6}$.  We assume the total number of samples
across all bins is large, hence the effect of considering $\lmax$ a
constant independent of the data is negligible. We then transform the
samples $\bml \to \bmx$ as
\begin{align}
  \lleq{gmf}
  x_j \equiv 1- \frac{\ell_j}{\lmax} > 0
\end{align}
and similarly for $L \to X$, then we infer $P(X \cond n, \bmx,
m)$. Only in the end, we transform back to a probability density in
$L$ via
\begin{align}
  \lleq{gmg}
  p(L  \cond n, \bml, m) = p(X  \cond n, \bmx, m) \frac{1}{\lmax}
\end{align}
We observe good agreement between the best fit of a Gamma distribution
to TARDIS samples; an example is shown shown in \reffig{gamma}.  While
not obvious at first, a major numerical simplification arises because
of the $L \to X$ transformation. Rewriting the master
equation~(\ref{pmp}) formally as
\begin{align}
  \lleq{gmh}
  p(L  \cond n, \bml, m) = \sum_{N} p(N \cond n) p(L \cond N, \bml, m)
\end{align}
we see that the right-hand side is a mixture density. In our
application, the distribution of a single luminosity, $p(L \cond N=1,
\bml, m)$ is narrow in the sense that the standard deviation relative
to the mean is at the percent level. This implies that $p(L \cond n,
\bml, m)$ has one mode from each $N$ and all modes are clearly
separated. The transformation $L \to x$ effects a strong overlap
between adjacent $p(X \cond N, \bmx, m)$, so $p(X \cond n, \bmx, m)$
can be well approximated by a unimodal distribution; see
\reffig{hedgehog}.  \fred{I'm not 100 \% sure but $p(X \cond n, \bmx,
  m)$ should still have $N \to \infty$ modes but they aren't visible}
\missingfigure{from hedgehog to single peak \label{fig:hedgehog}}

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.48\textwidth]{gamma}
  \fred{Plot with more samples to see dip at $x=0$.}
  \caption{Gamma distribution fit. }
  \label{fig:gamma}
\end{figure}

@@

The following approximations can be used in different regions of the
parameter space spanned by $n$ and $m$.

\begin{mnumerate}
  \item \label{large-m} For $m$ large, the posterior $p(\bmphi \cond m, \bmx) =
  p(\bmx\cond m,\bmphi) \, p(\bmphi) / p(\bmx \cond m)$ approaches a
  multivariate Gaussian. We can then compute the various moments
  $E_\bmphi[f(x)]$ with $f(x)=1,x,x^2$ by Laplace's method to good
  accuracy.
  \item \label{very-large-m} In the limit $m \to \infty$, the covariance of the posterior
  $p(\bmphi \cond m, \bmx)$ tends to zero as $1/m$ and we can then fix
  the parameters $\bmphi$ at the posterior mode $\bmphi^*$; i.e.,
  $p(\bmphi \cond m, \bmx) \to \delta(\bmphi - \bmphi^*)$.
  \item \label{large-n} For $n$ large, the cumulative of the Negative
  Binomial $p(N \cond n)$ can be approximated by a Gaussian, defining
  a region around $N=n$ with the dominant contribution to the sum over
  $N$ in \refeq{gmd}. The contribution around $N=0$ is negligible, and
  we can replace $\sum_n \to \int dN$. The density $p(N \cond n) P(X
  \cond N,m, \bmx)$ is close to a Gaussian and the integral can be
  accurately approximated by Laplace's method which requires the
  Hessian and the mode. We benefit from partial error cancelation
  because we apply Laplace's method both in the numerator and
  denominator~\cite{tierney_accurate_1986} as each moment has to be
  normalized by the evidence $p(\bmx \cond m)$. The former is
  available in closed form and the latter has to be determined
  numerically. Fortunately, Newton's method leads to a solution in
  only a few steps~\cite{minka2002estimating}. \fred{The reference is
    for max. likelihood of a $\GammaDist(\cdot)$ but I'm sure the
    performance is similar for max. posterior and the first moments}
  \item \label{central-limit} The central limit theorem states that for large $N$,
  $P(X \cond N, \bmphi) \approx \mathcal{N}(X \cond N \,
  E_\bmphi[X_j], N \, V_\bmphi[X_j])$. For $m$ large enough, we could
  then plug in the sample-mean and sample-variance estimates and
  bypass the assumption of a Gamma distribution altogether.
  \item \label{large-n-large-m} For large $n$ and $m$, $ p(X\cond
  n,\bmx,m)$ can be approximated by a Gaussian whose mean and variance
  are given by \refeq{mgo} and \refeq{mgr} in terms of
  $\bmphi$-expectation values. The latter are integrals with no known
  closed-form solutions but we can easily and accurately approximate
  them with Laplace's method.
\end{mnumerate}
To simplify, we assume that $m$ is large. In our application, we may
have to combine several frequency bins to achieve $m \gtrsim 150$ for
the purpose of infering $\bmphi$. This is equivalent to assuming that
the distribution changes only slowly with the frequency, a condition
that we verified for TARDIS. It is thus justified to use
approximations \ref{large-m} or even \ref{very-large-m}.

>>>>

\usepackage[colorinlistoftodos]{todonotes} % Use 'disable' to remove todos in final version
\newcommand{\fred}[1]{\todo[color=orange!40,inline]{#1}} %
\newcommand{\checked}{\todo[color=green,noline]{\checkmark}} %
\newcommand{\checkedl}{\todo[color=brown]{\checkmark}} %
\newcommand{\hans}[1]{\todo[color=yellow!30,inline]{#1}} %
\newcommand{\hanslong}[2][]{\todo[%bordercolor=red,
  color=white,inline,caption={2do}, #1]{
    \begin{minipage}{\textwidth} #2\end{minipage}}} %

%% CHANGE THIS DATE AS YOU MODIFY THE FILE
\newcommand{\vdate}{160228}
\pagestyle{myheadings}
\markboth{Version \vdate} {Version \vdate}
\parindent=0mm
\def\pp{\hfill\par \vspace{\baselineskip}}
%% \def\pp{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{bm,amssymb,amsfonts,amsmath}  % bold math symbols, AMStex
\usepackage{graphicx}    % standard graphics
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\lleq}[1]{\label{#1} }
% TO REMOVE THE LABEL LETTERS FROM THE EQUATION DISPLAY, COMMENT OUT THIS LINE:
\renewcommand{\lleq}[1]{\label{#1} {\scriptstyle {\rm (#1)}} \hspace*{2ex} }
% \renewcommand{\labelenumi}{\arabic{section}.\arabic{subsection}.\arabic{enumi}}
% \renewcommand{\thesubsubsection}{\arabic{subsubsection}}

@@

\DeclareMathOperator{\GammaDist}{Gamma}
\newcommand{\npack}{n_p}
\newcommand{\lmax}{\ell_{\rm max}}
\newcommand{\Lmax}{L_{\rm max}}

@@

\textbf{Abstract} TODO\\

\hans{Temporary addition of TOC to make it easier to find stuff. Eventually it should
disappear again.}

\tableofcontents

\newpage %
\hanslong{ %
  Hans 160228: \textbf{Summary of issues, based on v160218}. \\

  Before getting into all sorts of detail, I wanted to understand the
  major changes and developments. To my mind they are:
  \begin{enumerate}
  \item Fred discovered that the rescaling $X = L/\Lmax$ may have
    unintended advantages. TARDIS and general supernova luminosity
    spectra are highly variable because it consists not only of smooth
    blackbody spectrum, but also of strong absorption lines (emission
    also?? see 2014 MNRAS Kerzendorf, Sim et al?). Since $\Lmax$
    varies exactly like the spectrum itself, all binwise
    individual-packet distributions are rescaled to the interval
    $[0,1]$ leaving only a far weaker change of \textit{shape} of the
    distribution rather than the \textit{scale}.

  \item Just below Eq. (bsi) and also in (pmda), we had made explicit
    that we would do all the calculations for each bin separately, so
    there would be lots of fit parameters. Fred's 160218 paragraph
    starting with ``In general, the distribution \ldots'' and Eq
    (pmdb) sees significant reduction in the number of parameters if
    the above remnant dependence of $p(\bmL\cond N,\bmphi)$ in (pmda)
    on shape could be captured in a slow-varying function
    $a(\nu,\bmphi)$.

  \item Separately, Fred introduces the idea of combining elementary
    $b$-bins into compound bins with index $m$ as this improves the
    statistics.

  \item The simplification for large $N$ etc permits a Gaussian
    approximation, hence Laplace, hence the key equation (pmp) may be
    simplified to some approximation, making use of the moment
    generating function eg Eq (mgd) unnecessary.

  \item In principle, the above sounds like a very good development.
    There are some important caveats:
    \begin{enumerate}
    \item I see no detail on how to find $a(\nu,\bmphi)$. The first
      question would be if there is some numerical indication for its
      form. This seems to be equivalent to asking what the dependence
      of (say) the Gamma parameter $\alpha$ on $\nu$ is, and that is
      of course of great interest anyway.

    \item My second comment is that $a(\nu,\bmphi)$ is probably just
      your normal parametrisation which enters best-fit algorithms.
      If we are to treat it as a best-fit parametrisation with
      metaparameters, then that may just bring back all the
      complications which we thought had disappeared. This would have
      to be investigated. If on the other hand you also want to
      treat $a(\nu,\bmphi)$ as some kind of ``constant function'',
      then one would have to set out how stable this constant function
      would be against various underlying changes. In other words, 
      if calculation of $a()$ is not part of the general algorithm,
      how is it fixed? and can it be fixed for all cases?

    \item I think there may be a problem of combining the two proposed
      simplifications $L \to X$ and $m$ compound bins. Depending on
      the variability of the parametrisation as function of $\nu$, the
      shape of $p(\bmL\cond N,\bmphi)$ for single bins with large
      $\Lmax$ may differ significantly from the shape for many small
      bins combined into one using $m$. In other words, rebinning
      and rescaling may not be compatible.
    \end{enumerate}

  \item Once these are cleared up, we can get into the details \ldots\\

  \end{enumerate}
} %

@@

\subsection{Definitions of relevant quantities}
\label{sec:dfnr}

\begin{mtemize}
\item Spectrum of luminosities in channels (or ``bins'')
  $b=0,1,\ldots,B$ is given by the telescope
\item In the simplest case, the frequency of channel $b$ is given by
  $\nu_b = \nu_0 + b\,\delta\nu$, where $\delta\nu$ is
  the bin width. In the general case, the bins are not of equal
  size. Let's then consider $\nu_b$ as the left edge of bin $b$,
  and $\nu_{B+1}$ is the right edge of the last bin; i.e., the
  maximum frequency.
  %%%
  \hans{The 2014 TARDIS paper I have seen uses wavelength $\lambda$,
    not frequency in their plots. Is Wolfgang's recent data all
    frequency data? If it is, ie if Wolfgang et al have decided to
    change to frequency, they should explain whether that will happen
    in their and our papers also. If they want to stick with
    wavelength, we should too. I think this is something Wolfgang
    should decide.

    If they say we should stay with wavelength, then I would suggest
    replacing $\nu$ by $\lambda$ here and throughout, and of course
    change the Poisson parameter below to something like $\nu$.}

  \item We treat the results of a single TARDIS run as \textit{data}
    %%%%%
    \fred{If in the end we want to compare to actual data from a
      telescope, I'd prefer to not call the TARDIS output
      ``data''. Why don't we just call it ``TARDIS output''?} %
    \hans{I agree, ``data'' is awkward, and it is indeed useful to
      keep data and $D$ for later. ``Output'' would be ok, except that
      we should change the symbol $D$ to something else too. Symbol
      $O$ is not good. For the moment I would suggest $D_S$ (subscript
      simulation) or $D_T$ since we know that $T$ will later stand for
      temperature. Another useful term would be ``simdata'', symbol
      $S$.}
    %%%%%
    $D$, i.e.\ fixed results which we do not question but accept as
    given.  This ``data'' consists of the total number of packets
    leaving the supernova $\npack$ plus the set of luminosities
    $\bml$ and frequencies $\bmnu$ of each packet,
  \begin{align}
    \lleq{dfc}
    D = \{ \npack, \bml, \bmnu \} \,,
  \end{align}
  where
  \begin{align}
    \lleq{dfd}
    % \bml_b &= (\ell_{b,1},\ell_{b,2},\ldots,\ell_{b,n_b})
    \dim \bml = \dim \bmnu = \npack \,.
  \end{align}
  In terms of the $\mathbf{1}_V(x)$ the indicator function requiring
  $x \in V$, the number of packets in bin $b$ is
  \begin{align}
    \lleq{dff}
    n_b = \sum_{j=1}^{\npack} \mathbf{1}_{[\nu_b,\nu_{b+1}]}( \nu_j )
  \end{align}
  and the total luminosity in bin $b$ is
  \begin{align}
    \lleq{dfe}
    \ell_b = \sum_{j=1}^{\npack}\ell_j \mathbf{1}_{[\nu_b,\nu_{b+1}]}( \nu_j )\,,
  \end{align}
  with $\npack = \sum_b n_b$ and the total supernova luminosity $\ell
  = \sum_b \ell_b$.

\end{mtemize}
\hans{Discussion of rebinning with index $m$?}

@@

While we treat the TARDIS output as fixed ``data'' as discussed above,
a repeat of the simulation with a different Monte Carlo seed would
evidently yield different ``data'' $D'$ even when all other simulation
parameters remain the same. Clearly, $n_b$ is a specific realization
of a stochastic variable $N_b$ governed by a model pdf $p(N_b)$. %
\fred{In our general case, $\npack = \sum_b n_b$ is an instance of a
  random variable because some packets leave the supernova, some
  don't. I guess in most applications the total number of events would
  be chosen and not random. We are fortunate that this difference does
  not matter because we consider bins independently. } %
\hans{I will come back to this later} %
Similarly, $\bml_b$ is a specific realization of a collection of
stochastic variables $\bmL_b =\{L_{b,1},\ldots,L_{b,\smN_b}\}$, where
the number $N_b$ of such variables is itself stochastic. The relevant
pdf for $\bmL_b$ therefore always presupposes knowledge of $N_b$, so
we must work with the conditional probability $p(\bmL_b\cond
N_b)$. This situation is captured mathematically as $N_b \sim p(N_b),
\bmL_b \sim p(\bmL_b\cond N_b)$ where $\sim$ means ``is by hypothesis
distributed in this model according to the pdf'', i.e.\
\begin{align*}
  (N_b,\bmL_b) &\ \sim\ p(N_b,\bmL_b\cond \bmphi_b,\lambda_b)
  = p(\bmL_b\cond N_b,\bmphi_b)\; p(N_b\cond \lambda_b),
\end{align*}
where $\bmphi_b$ is the set of parameters of the model pdf which
govern $\bmL_b$ and $\lambda_b$ is the parameter of the model pdf
governing $N_b$.  To simulate $(N_b,\bmL_b)$, we clearly first need to
generate one $N_b$, followed by the generation of $N$ packet
luminosities $\bmL_b$. This situation is termed a \textit{compound
  process}.

@@

\begin{mtemize}
\item We omit the dependence on the physical supernova parameters
  $\bmtheta$ from the notation below and instead condition on the
  TARDIS output $\npack, \bml, \bmnu$ directly.

\item In the remainder of this paper, we assume that the packet
  multiplicities $N_b$ and total luminosities $\bmL_b$ in any given
  bin $b$ are statistically independent of those in other bins. %
  \fred{We could motivate this in the usual way as the Poisson limit
    of the Binomial. The subtlety would then the be the random total
    number of packets (leaving the supernova).} %
  \hans{The limit would be from the multinomial with $B$ bins and
    $\npack$ packtes to product of Poissonians. One could put it in,
    but in my opinion it's not necessary. $B$ is large, so the
    constraint plays no role.} %
  This means that the joint probability over all bins factorises into
  individual probabilities for each bin,
  \begin{align}
    \lleq{bsi}
    p(N_0,\bmL_0\ldots,N_b,\bmL_b,\ldots,N_\smB,\bmL_\smB
    \cond \npack, \bml, \bmnu,\bmphi,\bmlambda,\hmod)
    &= \prod_{b=0}^B p(N_b,\bmL_b\cond n_b, \bml, \bmnu,\bmphi,\bmlambda,\hmod)
    % p(N_0,\ldots,N_b,\ldots,N_\smB \cond \bmtheta,\bmphi,\hmod)
    % &= \prod_{b=0}^B p(N_b\cond \bmtheta,\bmphi,\hmod)\\
    % p(L_0,\ldots,L_b,\ldots,L_\smB \cond N_0,\ldots,N_\smB,\bmtheta,\bmphi,\hmod)
    % &= \prod_{b=0}^B p(L_b\cond N_b,\bmtheta,\bmphi,\hmod)
  \end{align}

@@

\item We choose $N$ to follow a Poisson distribution \fred{this choice
    linked to limit of Multinomial, see above} with parameter
  $\lambda$ %
  \hans{I would change the Poisson parameter to $\nu$, see above comment.}
  \begin{align}
    \lleq{pmc}
    p(N\cond\lambda) &= \frac{e^{-\lambda}\lambda^N}{N!}
    \qquad N = 0,1,\ldots,\infty,\quad \lambda>0
  \end{align}

@@

  \item In general, the distribution of $L_j$ could be a of a
  different functional form in each bin. For simplicity, we assume it
  is identical but may depend on parameters that differ from bin to
  bin. In practice, the distribution of packets does not change
  abruptly with frequency. We therefore assume that any parameter
  needed to fully specify $p(L_j)$ is given as a continuous function
  of frequency and parameters $\bmphi$; i.e.,
  \begin{align}
    \lleq{pmdb}
    p(L_j) = p(L_j \cond \nu, \bmphi)
  \end{align}
  For example, assume $p(L_j)$ depends on one parameter $a$. If we fit
  separately in each bin, we would have $B+1$ fits and $B+1$
  parameters. In each such fit, there would be only few samples to
  constrain $a$ in each bin. Then instead of fitting $a$ separately in
  each bin, we assume $a = a(\nu, \bmphi)$. This greatly reduces the
  number of parameters if $B$ is large because a handful of
  parameters, $\bmphi$, can be inferred from the complete set of
  packets. Necessarily, $a$ is best determined for the frequency
  around which the density of packets is highest. %
  \hans{I would replace the word ``sample'' above and throughout the
    text by ``packet''. The word ``sample'' refers in statistics to
    the entire data set $D$, not just to one data point.} %

@@

\item Priors: metaparameter $a$ \fred{?}  %
  \hans{This is a placeholder to remind us that metaparameters need to
    be discussed at some point. Perhaps not here, but we should
    remember.}
\end{mtemize}

@@

the present case is also the Laplace transform. %
\fred{On Wikipedia, the MGF is defined as $\int_0^\infty dL\,p(L\cond
  n,\bml)\,e^{+tL}$. Then there are less minus signs to worry about} %
\hans{Yes, most books seem to use the plus sign. It makes use of
  Laplace transform harder, but I am happy to use plus.} %

@@

The Gamma distribution  with shape parameter $\alpha$ and rate
parameter $\beta$ is defined as
\begin{align}
  \lleq{gma}
  \GammaDist(x | \alpha, \beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} x^{\alpha-1} e^{-\beta x} .
\end{align}
The support of $\GammaDist(x \cond \alpha, \beta)$ is $[0, \infty]$
whereas the luminosity of real packets is bounded from above and has a
peak close to but not at the boundary. \fred{refer to plot of real
  luminosity packets in some frequency bin} In order to describe the
luminosity distribution of real packets through $\GammaDist(\alpha,
\beta)$, we have to transform the luminosities. To this end, we first
take the largest luminosity from the entire frequency range and
increase it slightly to avoid numerical difficulties at the end point
\begin{align}
  \lleq{gme}
  \lmax \equiv (1+\epsilon) \times \max_j \ell_j,
\end{align}
where $ \epsilon=10^{-6}$.  We assume the total number of samples
across all bins is large, hence the effect of considering $\lmax$ a
constant independent of the data is negligible. We then transform the
samples $\bml \to \bmx$ as
\begin{align}
  \lleq{gmf}
  x_j \equiv 1- \frac{\ell_j}{\lmax} > 0
\end{align}
and similarly for $L \to X$, then we infer $P(X \cond n, \bmx,
m)$. Only in the end, we transform back to a probability density in
$L$ via
\begin{align}
  \lleq{gmg}
  p(L  \cond n, \bml, m) = p(X  \cond n, \bmx, m) \frac{1}{\lmax}
\end{align}
\hans{So far, $m$ appears only in your todo comment above. I would suggest
  a short discussion of this already in Section \ref{sec:dfnr}. We
  could then simply rewrite all equations below in terms of $m$ rather
  than $b$.} %
We observe good agreement between the best fit of a Gamma distribution
to TARDIS samples; an example is shown shown in \reffig{gamma}.  While
not obvious at first, a major numerical simplification arises because
of the $L \to X$ transformation. Rewriting the master
equation~(\ref{pmp}) formally as
\begin{align}
  \lleq{gmh}
  p(L  \cond n, \bml, m) = \sum_{N} p(N \cond n) p(L \cond N, \bml, m)
\end{align}
we see that the right-hand side is a mixture density. In our
application, the distribution of a single luminosity, $p(L \cond N=1,
\bml, m)$ is narrow in the sense that the standard deviation relative
to the mean is at the percent level. This implies that $p(L \cond n,
\bml, m)$ has one mode from each $N$ and all modes are clearly
separated. The transformation $L \to x$ effects a strong overlap
between adjacent $p(X \cond N, \bmx, m)$, so $p(X \cond n, \bmx, m)$
can be well approximated by a unimodal distribution; see
\reffig{hedgehog}.  \fred{I'm not 100 \% sure but $p(X \cond n, \bmx,
  m)$ should still have $N \to \infty$ modes but they aren't visible}
\missingfigure{from hedgehog to single peak \label{fig:hedgehog}}

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.48\textwidth]{gamma}
  \fred{Plot with more samples to see dip at $x=0$.} %
  \hans{I assume the KDE fit will disappear from the paper later as it
    overshoots the zero point.}
  \caption{Gamma distribution fit. }
  \label{fig:gamma}
\end{figure}

@@

The following approximations can be used in different regions of the
parameter space spanned by $n$ and $m$. %
\hans{$n$ is TARDIS output ``data'', so not a parameter. From the
  above paragraph, $m$ seems to be a parameter governing rebinning to
  nonuniform bins. I would have no idea how to handle $m$ formally in
  a Bayesian way. If $m{=}150$ is the result of optimising the fitting
  stability, then presumably it will be kept constant. In that case,
  it would be treated like a constant rather than a parameter in the
  usual sense.}
\begin{mnumerate}
  \item \label{large-m} For $m$ large, the posterior $p(\bmphi \cond m, \bmx) =
  p(\bmx\cond m,\bmphi) \, p(\bmphi) / p(\bmx \cond m)$ approaches a
  multivariate Gaussian. We can then compute the various moments
  $E_\bmphi[f(x)]$ with $f(x)=1,x,x^2$ by Laplace's method to good
  accuracy.
  \item \label{very-large-m} In the limit $m \to \infty$, the covariance of the posterior
  $p(\bmphi \cond m, \bmx)$ tends to zero as $1/m$ and we can then fix
  the parameters $\bmphi$ at the posterior mode $\bmphi^*$; i.e.,
  $p(\bmphi \cond m, \bmx) \to \delta(\bmphi - \bmphi^*)$.
  \item \label{large-n} For $n$ large, the cumulative of the Negative
  Binomial $p(N \cond n)$ can be approximated by a Gaussian, defining
  a region around $N=n$ with the dominant contribution to the sum over
  $N$ in \refeq{gmd}. The contribution around $N=0$ is negligible, and
  we can replace $\sum_n \to \int dN$. The density $p(N \cond n) P(X
  \cond N,m, \bmx)$ is close to a Gaussian and the integral can be
  accurately approximated by Laplace's method which requires the
  Hessian and the mode. We benefit from partial error cancelation
  because we apply Laplace's method both in the numerator and
  denominator~\cite{tierney_accurate_1986} as each moment has to be
  normalized by the evidence $p(\bmx \cond m)$. The former is
  available in closed form and the latter has to be determined
  numerically. Fortunately, Newton's method leads to a solution in
  only a few steps~\cite{minka2002estimating}. \fred{The reference is
    for max. likelihood of a $\GammaDist(\cdot)$ but I'm sure the
    performance is similar for max. posterior and the first moments}
  \item \label{central-limit} The central limit theorem states that for large $N$,
  $P(X \cond N, \bmphi) \approx \mathcal{N}(X \cond N \,
  E_\bmphi[X_j], N \, V_\bmphi[X_j])$. For $m$ large enough, we could
  then plug in the sample-mean and sample-variance estimates and
  bypass the assumption of a Gamma distribution altogether.
  \item \label{large-n-large-m} For large $n$ and $m$, $ p(X\cond
  n,\bmx,m)$ can be approximated by a Gaussian whose mean and variance
  are given by \refeq{mgo} and \refeq{mgr} in terms of
  $\bmphi$-expectation values. The latter are integrals with no known
  closed-form solutions but we can easily and accurately approximate
  them with Laplace's method.
\end{mnumerate}
To simplify, we assume that $m$ is large. In our application, we may
have to combine several frequency bins to achieve $m \gtrsim 150$ for
the purpose of infering $\bmphi$. This is equivalent to assuming that
the distribution changes only slowly with the frequency, a condition
that we verified for TARDIS. It is thus justified to use
approximations \ref{large-m} or even \ref{very-large-m}.
